\Title

### 2021-01-06 manuscript/manuscript.tex
Standardize use of "Equivalent sources" and "gradient-boosted"

% Define title, authors, affiliations and DOI
% ===========================================
\newcommand{\Title}{
    Gradient boosted equivalent sources
}
\newcommand{\Author}{
    S.R. Soler,
    L. Uieda
}
\newcommand{\AuthorAffil}{
    {\large
        Santiago R. Soler$^{1,2}$,
        Leonardo Uieda$^{3}$
    }
    \\[0.4cm]
    {\small $^{1}$CONICET, Argentina (santiago.r.soler@gmail.com)} \\
    {\small $^{2}$Instituto Geofísico Sismológico Volponi, UNSJ, Argentina} \\
    {\small $^{3}$Department of Earth, Ocean and Ecological Sciences, School of Environmental Sciences, University of Liverpool, UK} \\
}
\newcommand{\DOI}{
    doi:\href{https://doi.org/xxx.xxx/xxxxxx}{xxx.xxx/xxxxxx}
}
\newcommand{\DOILink}{
    \href{https://doi.org/xxx.xxx/xxxxxx}{doi.org/xxx.xxx/xxxxxx}
}

@@

A widely used method for interpolating gravity and magnetic data
is the equivalent source technique (also known as equivalent layer, radial
basis functions, or Green's functions interpolation), first introduced by
\citet{dampney1969}.
It consists in fitting a model of elementary sources to the data and using this
model to predict new data values.
Besides interpolation, equivalent sources have been used for
reduction-to-the-pole of magnetic data
\citep{silva1986, nakatsuka2006, guspi2009}, upward
continuation \citep{emilia1973, li2010}, joint processing of gravity gradient
data \citep{barnes2011}, modelling the lithospheric magnetic field
\citep{kother2015}, recovering the magnetic induction vector from
total-field magnetic anomalies \citep{li2020}, and more.

Many variants of the equivalent source technique have been proposed, often
attempting to obtain faster or more accurate solutions.
The key factors that vary between them are: (i) the type of source, (ii)
the location of the sources, and (iii) the solution strategy.
The type of source is most commonly a point mass for gravity or dipole for
magnetics \citep[e.g.,~][]{vonfrese1981, silva1986, mendonca1994, siqueira2017}.
However, right-rectangular prisms \citep[e.g.,][]{barnes2011, jirigalatu2019,
li2020} and tesseroids \citep{bouman2016} have also been used successfully.
In fact, even point sources with a simple inverse distance function, instead of
actual gravity or magnetic fields, can be used as
equivalent sources \citep{cordell1992}.
The sources are often distributed on a regular grid at a constant depth
\citep[e.g.,~][]{leao1989, barnes2011, oliveira2013}
or placed beneath each data point \citep[e.g.,~][]{cordell1992, siqueira2017}.
The model is usually estimated through damped least-squares, which imposes a
heavy computational load when the number of data points is large (e.g.,
airborne and satellite surveys).
To reduce the computational load, \citet{mendonca1994} built the solution
iteratively by incorporating one data point at a time using the ``equivalent
data concept''.
\citet{leao1989} processed the input data using a moving window, only fitting the
data inside the window and predicting observations at the center of the window.
\citet{li2010} and \citet{barnes2011} apply different operations to generate a
sparse representation of the sensitive matrix (respectively, wavelet
compression and quadtree discretization), which significantly improves the
speed of the least-squares solution.
\citet{oliveira2013} parametrized the equivalent layer as a piecewise bivariate
polynomial function, reducing the number of parameters in the solution.
\citet{siqueira2017} developed an iterative solution in which the sensitivity
matrix is transformed into a diagonal matrix with constant terms through the
``excess mass criterion''.
\citet{jirigalatu2019} applied the Gauss-FFT method to speed up the forward
modelling operations and solved the least-squares problem using steepest
descent to avoid calculating the Hessian matrix and solving linear systems.

@@

The $\mathbf{A}_k$ matrices have only $N \times M_k$ elements
(where $M_k$ is the number sources on the $k-$th window), which can be
considerably smaller than the $N \times M$ elements of $\mathbf{A}$.
Therefore, the gradient boosting method allows us to fit
equivalent sources models that would produce Jacobian matrices that are larger
than the available computer memory.
Furthermore, we can increase or decrease the size of the overlapping windows as
needed depending on the number of sources in the model and the available
computer memory.

We can improve the efficiency of the algorithm further by using only the $N_k$ data
points that fall within the $k$-th window for fitting the sources
(steps \ref{alg:scale} and \ref{alg:fit} of
algorithm~\ref{alg:gradient_boosting}),
while using all available data when calculating the step size and residuals
(steps \ref{alg:predicted}, \ref{alg:stepsize}, and \ref{alg:residual}
of algorithm~\ref{alg:gradient_boosting}).
By doing so, we can reduce the size of the Jacobian matrix further to $N_k
\times M_k$ elements.
The forward modeling operation performed in step \ref{alg:predicted} can be
done by a summation (equation~\ref{eq:forward-sum})
instead of a matrix-vector product, which allows us to avoid computing and
storing the larger $N \times M_k$ Jacobian at any point.
Algorithm~\ref{alg:gradient_boosting_window} incorporates these changes and is
the final \textit{gradient boosted equivalent source algorithm}.

@@

    $\mathbf{r}_k = \mathbf{r}_{k - 1} - \alpha_k\mathbf{d}_k$
    \;
  }
  \BlankLine
  \setstretch{1}
  \caption{ Gradient boosted equivalent source algorithm. }
  \label{alg:gradient_boosting_window}
\end{algorithm}

@@

\section{Performance of gradient boosted equivalent sources}

>>>>

% Define title, authors, affiliations and DOI
% ===========================================
\newcommand{\Title}{
    Gradient-boosted equivalent sources
}
\newcommand{\Author}{
    S.R. Soler,
    L. Uieda
}
\newcommand{\AuthorAffil}{
    {\large
        Santiago R. Soler$^{1,2}$,
        Leonardo Uieda$^{3}$
    }
    \\[0.4cm]
    {\small $^{1}$CONICET, Argentina (santiago.r.soler@gmail.com)} \\
    {\small $^{2}$Instituto Geofísico Sismológico Volponi, UNSJ, Argentina} \\
    {\small $^{3}$Department of Earth, Ocean and Ecological Sciences, School of Environmental Sciences, University of Liverpool, UK} \\
}
\newcommand{\DOI}{
    doi:\href{https://doi.org/xxx.xxx/xxxxxx}{xxx.xxx/xxxxxx}
}
\newcommand{\DOILink}{
    \href{https://doi.org/xxx.xxx/xxxxxx}{doi.org/xxx.xxx/xxxxxx}
}

@@

A widely used method for interpolating gravity and magnetic data
is the equivalent sources technique (also known as equivalent layer, radial
basis functions, or Green's functions interpolation), first introduced by
\citet{dampney1969}.
It consists in fitting a model of elementary sources to the data and using this
model to predict new data values.
Besides interpolation, equivalent sources have been used for
reduction-to-the-pole of magnetic data
\citep{silva1986, nakatsuka2006, guspi2009}, upward
continuation \citep{emilia1973, li2010}, joint processing of gravity gradient
data \citep{barnes2011}, modelling the lithospheric magnetic field
\citep{kother2015}, recovering the magnetic induction vector from
total-field magnetic anomalies \citep{li2020}, and more.

Many variants of the equivalent sources technique have been proposed, often
attempting to obtain faster or more accurate solutions.
The key factors that vary between them are: (i) the type of source, (ii)
the location of the sources, and (iii) the solution strategy.
The type of source is most commonly a point mass for gravity or dipole for
magnetics \citep[e.g.,~][]{vonfrese1981, silva1986, mendonca1994, siqueira2017}.
However, right-rectangular prisms \citep[e.g.,][]{barnes2011, jirigalatu2019,
li2020} and tesseroids \citep{bouman2016} have also been used successfully.
In fact, even point sources with a simple inverse distance function, instead of
actual gravity or magnetic fields, can be used as
equivalent sources \citep{cordell1992}.
The sources are often distributed on a regular grid at a constant depth
\citep[e.g.,~][]{leao1989, barnes2011, oliveira2013}
or placed beneath each data point \citep[e.g.,~][]{cordell1992, siqueira2017}.
The model is usually estimated through damped least-squares, which imposes a
heavy computational load when the number of data points is large (e.g.,
airborne and satellite surveys).
To reduce the computational load, \citet{mendonca1994} built the solution
iteratively by incorporating one data point at a time using the ``equivalent
data concept''.
\citet{leao1989} processed the input data using a moving window, only fitting the
data inside the window and predicting observations at the center of the window.
\citet{li2010} and \citet{barnes2011} apply different operations to generate a
sparse representation of the sensitive matrix (respectively, wavelet
compression and quadtree discretization), which significantly improves the
speed of the least-squares solution.
\citet{oliveira2013} parametrized the equivalent layer as a piecewise bivariate
polynomial function, reducing the number of parameters in the solution.
\citet{siqueira2017} developed an iterative solution in which the sensitivity
matrix is transformed into a diagonal matrix with constant terms through the
``excess mass criterion''.
\citet{jirigalatu2019} applied the Gauss-FFT method to speed up the forward
modelling operations and solved the least-squares problem using steepest
descent to avoid calculating the Hessian matrix and solving linear systems.

@@

The $\mathbf{A}_k$ matrices have only $N \times M_k$ elements
(where $M_k$ is the number sources on the $k-$th window), which can be
considerably smaller than the $N \times M$ elements of $\mathbf{A}$.
Therefore, the gradient boosting method allows us to fit
equivalent source models that would produce Jacobian matrices that are larger
than the available computer memory.
Furthermore, we can increase or decrease the size of the overlapping windows as
needed depending on the number of sources in the model and the available
computer memory.

We can improve the efficiency of the algorithm further by using only the $N_k$ data
points that fall within the $k$-th window for fitting the sources
(steps \ref{alg:scale} and \ref{alg:fit} of
algorithm~\ref{alg:gradient_boosting}),
while using all available data when calculating the step size and residuals
(steps \ref{alg:predicted}, \ref{alg:stepsize}, and \ref{alg:residual}
of algorithm~\ref{alg:gradient_boosting}).
By doing so, we can reduce the size of the Jacobian matrix further to $N_k
\times M_k$ elements.
The forward modeling operation performed in step \ref{alg:predicted} can be
done by a summation (equation~\ref{eq:forward-sum})
instead of a matrix-vector product, which allows us to avoid computing and
storing the larger $N \times M_k$ Jacobian at any point.
Algorithm~\ref{alg:gradient_boosting_window} incorporates these changes and is
the final \textit{gradient-boosted equivalent source algorithm}.

@@

    $\mathbf{r}_k = \mathbf{r}_{k - 1} - \alpha_k\mathbf{d}_k$
    \;
  }
  \BlankLine
  \setstretch{1}
  \caption{Gradient-boosted equivalent source algorithm.}
  \label{alg:gradient_boosting_window}
\end{algorithm}

@@

\section{Performance of gradient-boosted equivalent sources}
